Bias in the Estimation of the Value of VRA over URA: A Case of Mis-speficied Geographically Weighted Regression Models

### 2023-11-15 Writing/manuscript_gwr.rmd
manuscript modified

#* read the results rds file
results <-
  here("Shared/Results/gaussian/pi_data.rds") %>%
  readRDS() %>%
  .[, type := ifelse(transfer == 0, "GWRR", "GWRT")] %>%
  .[, bias := pi_diff_est - pi_diff] %>%
  # only display the 25% (5.44), 50% (6.56), 75% (7.67) price ratio
  .[pRatio %in% c(5.44, 6.56, 7.67), ] %>%
  # === label price ratio
  .[, pLabelName := "Price Ratio (N/corn)"] %>%
  .[pRatio == 5.44, pLabel := "Low (5.44)"] %>%
  .[pRatio == 6.56, pLabel := "Middle (6.56)"] %>%
  .[pRatio == 7.67, pLabel := "High (7.67)"] %>%
  .[, pLabel := factor(pLabel, levels = c("Low (5.44)", "Middle (6.56)", "High (7.67)"))]

@@

**Abstract**: Geographically weighted regression (GWR) has been presented as a valuable tool for the estimation of site-specific yield response functions used to derive site-specific input rate recommendations from data generated by on-farm precision experimentation (OFPE). But most of the applications of GWR in this context have assumed a quadratic yield response function, despite a scientific consensus that the yield-input relationships modeled are better represented by the quadratic-plateau functional form. The key objective of this paper is to investigate whether, when the true functional form of yield response is quadratic-plateau, assuming a quadratic form introduces bias into the GWR estimation of the economic value of site-specific recommendation maps. We consider two sets of functional-form specifications used by recently published studies of site-specific input management that apply GWR to data from OFPE. We show that both GWR approaches suffer significant positive bias, which can lead to large over-estimation of the economic value of applying GWR methods. We conclude that in many circumstances the information gleaned from using GWR may be far less valuable than has been claimed.  We encourage the statistical community to develop tools in software packages providing GWR that allow more flexibility in functional form assumptions.

@@

GWR has been presented as a valuable tool for the estimation of site-specific yield response functions used to derive site-specific input rate recommendations from data generated by OFPE [@rakshit2020; @trevisan2021spatial]. @trevisan2021spatial, in particular, reported an unusually high potential gain of $58 per hectare from transitioning from the optimal uniform application rate to optimal site-specific application rates. Such a figure is notably higher than what has been suggested by previous studies on the value of variable rate application (VRA) over uniform rate application (URA) [@bullock1998does; @koch2004economic; @thrikawala1999economic]. We suspect that their reported estimate for the benefit of GWR-based VRA over URA may be significantly over-estimated for the very reasons mentioned earlier. First, their model for corn yield response to nitrogen is mostly likely mis-specified. Numerous past agronomic studies recognize that corn yield follows a quadratic-plateau response to nitrogen, yet a simple quadratic model was employed by @trevisan2021spatial. Similarly, @rakshit2020 also used quadratic model. Ideally, the GWR model should be specified with quadratic-plateau or more flexible functional forms. However, no software package is readily usable to estimate GWR with quadratic-plateau fucntional form. Second, GWR itself has faced criticism regarding the statistical reliability of its coefficient estimations [@griffith2008spatial; @harris2019simulation; @paez2011simulation; @da2016multiple]. We hypothesize that the intersection of these two issues—the model mis-specification and the inherent limitations of GWR—may have led to a substantial overestimation of the economic value of switching from URA to VRA.

(Taro: I feel the primary objective should be two-folded. As Nathan summarized in his email, they should be (1) the true economic value of GWR-based VRA and, (2) the bias in the estimation of that value. But, you make the final call.) The primary objective of this study is to examine the degree and direction of bias in the estimation of the economic value of GWR-based variable nitrogen rate recommendations, especially when the yield response to nitrogen is incorrectly specified as quadratic in the GWR model instead of the correct quadratic-plateau function. (Nathan's emphasis: "a difference between quadratic GWR and a more flexible functional form with no spatial heterogeneity". Shall we put it here?)  We will employ Monte Carlo simulations, which allow us to accurately define and know the true yield response function within our data-generating process, providing a means to quantify the degree of bias present.

Differing from the works of @griffith2008spatial, @harris2019simulation, @paez2011simulation, and @da2016multiple, our goal is not to critique the bias in GWR's coefficient estimates per se. Instead, we seek to assess the economic repercussions^[Note that while the environmental consequences of fertilizer management carry significant societal importance, they fall outside the purview of this study. Herein, "economically optimal" pertains strictly to the maximization of private production profitability for farmers and does not imply "socially optimal."] stemming from functional form mis-specifcatin combined with statistical inaccuracies in the coefficient estimation of the model. We found that a mis-specified GWR model tends to result in substantial over-estimation of the economic value of GWR-based VRA over URA.

@@

In practice, since the true responses $f_{j, i}(\cdot)$ are unknown, the calculation of the economic value of GWR-based VRA is always based on assumption that the GWR-estimated yield response functions are true. In this study, the estimated profits of VRA (based on GWR) and URA (based on SCAM) based on the GWR-estimated responses were:

@@

## Comparing the values of GWR-based VRA and SCAM-based URA

Figure \@ref(fig:pi-dif-dist) shows the distribution of $\Delta^{GWR}$, the true profit difference between the GWR-based VRA and the SCAM-based URA, at three different price scenarios which represent the historically low (5.44), middle (6.56), and high (7.67) nitrogen-corn price ratio (\$/kg : \$/kg). Historical corn ($p$) and nitrogen fertilizer ($w$) prices from 2002 to 2022 were obtained from the USDA National Agricultural Statistics Service [@NASS2022] and DTN Retail Fertilizer Trends [@DTN2022], and were transformed into monthly data. The nitrogen-corn price ratios were calculated as $w/p$. Further information regarding the collection, calculation, and summary statistics of the price data can be found in the supplementary appendix (Appendix 1: Historical Corn and Nitrogen Prices).

The comparative advantage of GWRR and GWRT varies significantly depending on the nitrogen-corn price ratio. Specifically, GWRR performs better as nitrogen becomes more expensive relative to corn, while the opposite is the case for GWRT. At the low price ratio of 5.44, VRA based on GWRR is no better than SCAM-based URA. However, when the price ratio increases to $6.56$, the median benefit of GWRR is $`r round(mean_data_value[type == "GWRR" & pLabel == "Middle", pi_diff], digits = 2)` per hectare. At the high price ratio of 7.67, the median benefit is $`r round(mean_data_value[type == "GWRR" & pLabel == "High", pi_diff], digits = 2)` per hectare. On the other hand, the median benefit of VRA based on GWRT at the low price ratio of 5.44 is $`r round(mean_data_value[type == "GWRT" & pLabel == "Low", pi_diff], digits = 2)` per hectare, but the benefit is reduced to $`r round(mean_data_value[type == "GWRT" & pLabel == "Middle", pi_diff], digits = 2)` per hectare when the price ratio is 6.56, and $`r round(mean_data_value[type == "GWRT" & pLabel == "High", pi_diff], digits = 2)` per hectare when 7.67.

Figure \@ref(fig:bias-est-eonr) provides insights into this phenomenon, showing the distribution of the average ratio of estimated EONR to true EONR. A value lower than 1 means that EONR is underestimated on average. GWRR tends to over-estimate EONR at lower price ratios, but the magnitude of over-estimation falls as the price ratio rises. The opposite trend is true for GWRT. At the high price ratio, the degree of EONR over-estimation is quite significant. Since a high-price ratio means nitrogen is relatively expensive, the economic cost of over-estimation of EONR increases with the price ratio.

To illustrate the difference in the GWRR and GWRT outcomes, Figure \@ref(fig:true-vs-estimated-optn-gwr-r) presents for a single simulation round the estimated site-specific EONR plotted against the true site-specific EONR. The site-specific EONR estimates are both inaccurate, but differently inaccurate. The range of EONR estimates by GWRR is very wide because the coefficient on $N^2$ is allowed to vary site specifically in GWRR. Since the estimation of the coefficient on $N^2$ is highly inaccurate, the range of estimated EONRs is much wider than the range of true site-specific EONRs. However, GWRT's range of EONR estimates is very narrow because the coefficient on $N^2$ is constrained to be equal across sites, making GWRT fail to capture the EONR's true spatial heterogeneity.

## Bias in the estimation of the value of GWR-based VRA

Next, we examine the bias in the estimated value of GWR-based VRA. Figure \@ref(fig:bias-est-pi) shows the distribution of the bias in the estimation of the value of GWR-based VRA ($\hat{\Delta}^{GWR} - \Delta^{GWR}$). For example, the value of 10 would mean that the estimated value of GWR-based VRA over-estimates its true value by $10 per ha. As the figure shows, GWRR-based VRA appears to perform better than it truly does, presenting significant positive bias across the relative price range. That is, researchers would tend to significantly over-estimate the value of VRA based on GWRR. When the GWR estimates are considered to define true yield response, the inaccuracy of GWRR generates over-estimation of the heterogeneity of yield response within the field, which can lead to significant over-estimation of the value of GWRR itself.

GWRT exhibits a different pattern of bias. At low price ratios, the GWRT approach displays smaller bias compared to GWRR. This is partially because GWRT tends to understate the heterogeneity of the site-specific yield response functions as was seen in Figure \@ref(fig:true-vs-estimated-optn-gwr-r). As the price ratio rises, bias increases. This is due to the significant over-estimation of EONR by GWRT, as illustrated by Figure \@ref(fig:bias-est-eonr). Unlike the GWRR approach, the distribution of the bias by the GWRT approach has a fat tail, indicating that there is a greater chance of encountering large positive bias in the estimated value of VRA over URA.

Figure \@ref(fig:why-bias-many) illustrates how over-estimation of the value of VRA can occur. It shows yield response functions estimated by GWRT and estimated yields associated with the GWR-based EONR rates and SCAM-based EONR rates for 50 randomly chosen cells in a single simulation. Since GWR-based EONRs were obtained assuming the yield response functions estimated by GWR, by construction they out-perform the SCAM-based estimates in every cell. For many cells, the SCAM-based estimated EONR is significantly higher than the actual yield-maximizing N rate, resulting in an estimated yield far lower because of the quadratic functional form assumption. Figure \@ref(fig:why-bias-single) panel (a) illustrates the true quadratic-plateau yield response and estimated yield response by the mis-specified quadratic GWR for a single cell. The cell's true yield response plateaus at a nitrogen rate of about $`r round(il_data_oe[aunit_id %in% single_aunit, Nk], digits = 0)`$ kg ha$^{-1}$, but the GWR-estimated yield response function suggests that yield should peak at a nitrogen rate of about $`r round(point_data_gwr[type == "GWR", N], digits = 0)`$ kg ha$^{-1}$ and decline thereafter. The true yield differential between recommended N rates by GWR and SCAM is 0, while the estimated yield differential is about `r round(point_data_gwr[type == "GWR", yield] - point_data_gwr[type == "SCAM", yield], digits = 2)` ton ha$^{-1}$, leading to the over-estimation of the GWR yield benefit. Panel (b) illustrates the corresponding true and estimated profit response curves. Based on the true profit response, applying the GWR-based recommendation is about \$`r round(true_yield_data[type != "True", diff(profit)], digits = 0)` ha$^{-1}$ less profitable than the SCAM-based recommendation, where this profit differential comes solely from the difference in nitrogen cost since GWR and SCAM yields are equal. However, when calculations are based on the cell's GWR-estimated response, profit from the GWR-based recommendation exceeds that from the SCAM-based recommendation by about \$`r round(- point_data_gwr[, diff(profit)], digits = 0)` ha$^{-1}$, which is much larger than the true value of \$`r round(true_yield_data[type != "True", diff(profit)], digits = 0)` ha$^{-1}$. Beyond the difference in nitrogen cost, the falsely estimated yield boost by GWR contributes to the profit differential, thus significantly over-estimating the actual value of GWR-based VRA.

@@

Our simulation results show that site-specific $R_x$s derived by the quadratic GWR models (GWRR and GWRT) has little economic value over URA derived by a correctly specified uniform response model (SCAM) when the true yield response function is quadratic-plateau. However, we showed that the estimated value of the VRA based on the quadratic GWR models were much higher than their actualy values. This finding reveals the severe bias in estimating the value of VRA over URA when the VRA was derived from quadratic GWR when the true yield response function follows the quadratic-plateau functional form, such as in the corn-nitrogen relationship. Precision agriculture practitioners should be cautious when using GWR for site-specific $R_x$s.

Our results clearly show that it can be problematic to use GWR under the assumption of a quadratic functional form when the true yield-input relationship is quadratic-plateau. @trevisan2021spatial had to choose the quadratic functional form because no readily usable statistical software or package could be used to implement quadratic-plateau GWR at the time. The only nonlinear model available at the moment is the linear-plateau GWR by @lambert2022geographically. Unfortunately, the estimation of the economically optimal input rate cannot be adjusted according to price using the linear-plateau model as the economically optimal input rate occurs always at the kink (critical input rate where yield plateau is reached). Given that many yield-input relationships follow quadratic-plateau functional form for many crops, it would be fruitful to develop a GWR method with quadratic-plateau functional form. It should be also possible and interesting to build semi-parametric GWR where the yield response functions are estimated without assuming particular functional forms (like SCAM). It is also noted that the recently developed multi-scale GWR model [@fotheringham2017multiscale; @yu2020inference; @comber2022route] allows the bandwidth to be derived separately for each covariate in the model, which would potentially improve the performances of GWR estimates and may need to be incorporated into future research. Note, however, that multi-scale GWR models are not necessary in our simulations as we have only one covariate ($N$) in our simulations.

Our results clearly show that it can be problematic to use GWR under the assumption of a quadratic functional form when the true yield-input relationship is quadratic-plateau. @trevisan2021spatial had to choose the quadratic functional form because no readily usable statistical software or package could be used to implement nonlinear GWR. The only nonlinear model available at the moment is the linear-plateau GWR by @lambert2022geographically. Unfortunately, the estimation of the economically optimal input rate cannot be adjusted according to price using the linear-plateau model as the economically optimal input rate occurs always at the kink (critical input rate where yield plateau is reached). Given that many yield-input relationships follow quadratic-plateau functional form for many crops, it would be fruitful to develop a GWR method with quadratic-plateau functional form. It should be also possible and interesting to build semi-parametric GWR where the yield response functions are estimated without assuming particular functional forms (like SCAM). It is also noted that the recently developed multi-scale GWR model [@fotheringham2017multiscale; @yu2020inference; @comber2022route] allows the bandwidth to be derived separately for each covariate in the model, which would potentially improve the performances of GWR estimates and may need to be incorporated into future research. Note, however, that multi-scale GWR models are not necessary in our simulations as we have only one covariate ($N$) in our simulations.

Most importantly, whether using GWRT or GWRR, researchers must be cautious about their estimated values of VRA when the model is assumed to be quadratic but the true underlying yield-input response is quadratic-plateau. Our results show that it is likely that they will over-estimate their true values and the bias can be misleadingly high with non-negligible probability. The extremely high values of VRA reported in @trevisan2021spatial are likely to have reflected this bias. Indeed, the bias problem we have raised here is likely to be observed for other statistical methods as well because the bias comes from the assumption that the _estimated_ yield response functions are true. Whatever the underlying statistical methods for yield response functions may be, researchers are likely to overestimate the true value of their methods. It would be an interesting future research to investigate the bias in other commonly used statistical methods in the literature on site-specific yield response estimation (such as regression with interaction terms of field characteristics, machine learning, and more) by using a similar simulation-based evaluation approach as done in this study.

<!-- # Conclusions

This article examined the economic value of GWR-based VRA and the bias in its estimation in the context of OFPE. We showed that VRA based on GWR provides only small economic benefits over URA based on SCAM. The relative performances of the model specifications used by @trevisan2021spatial and @rakshit2020 differs based on nitrogen-corn price ratio. Finally and most important, we found that both specifications can result in significant over-estimation of the true value of GWR-based site-specific nitrogen rate recommendations. It can be concluded that researchers need to be cautious about promoting the GWR-based approach to VRA recommendation. -->

@@

```{r, fig.id = "pi-dif-dist", fig.cap = "The true economic value of GWR-based VRA over SCAM-based URA by two GWR models (GWRR and GWRT) and at three nitrogen-corn price ratios", out.width = "6.5in", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_value.png"))
g_value
```

@@

```{r fig.id = "why-bias-many", fig.cap = "The cause of significant over-estimation of the value of GWR-based VRA", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_why_bias_many.png"))
g_why_bias_many
```

```{r fig.id = "why-bias-single", fig.cap = "An illustration of over-estimation of the value of GWR-based VRA over SCAM-based URA",fig.width = 6.5, fig.height = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_why_bias_single.png"))
g_why_bias_single
```

>>>>

#* read the results rds file
results <-
    here("Shared/Results/gaussian/pi_data.rds") %>%
    readRDS() %>%
    # === do not need QD result
    .[model!="QD", ] %>% 
    # === modify model names
    .[model=="GWR-R", model := "GWRR"] %>% 
    .[model=="GWR-T", model := "GWRT"] %>% 
    # === only need  25% (5.44), 50% (6.56), 75% (7.67) price ratio
    .[pRatio %in% c(5.44, 6.56, 7.67), ] %>%
    # === label price ratio
    .[pRatio == 5.44, pLabel := "5.44 (Low)"] %>%
    .[pRatio == 6.56, pLabel := "6.56 (Middle)"] %>%
    .[pRatio == 7.67, pLabel := "7.67 (High)"] %>%
    # === estimation bias
    .[, bias := pi_diff_est - pi_diff]

@@

**Abstract**: Geographically weighted regression (GWR) has been presented as a valuable tool for the estimation of site-specific yield response functions used to derive site-specific input rate recommendations from data generated by on-farm precision experimentation (OFPE). But most of the applications of GWR in this context have assumed a quadratic yield response function, despite a scientific consensus that the yield-input relationships modeled are better represented by the quadratic-plateau functional form. The key objective of this paper is to investigate whether, when the true functional form of yield response is quadratic-plateau, assuming a quadratic form introduces bias into the GWR estimation of the economic value of site-specific recommendation maps. We consider two sets of functional-form specifications used by recently published studies of site-specific input management that apply GWR to data from OFPE. We show that both GWR approaches suffer significant positive bias, which can lead to large overestimation of the economic value of applying GWR methods. We conclude that in many circumstances the information gleaned from using GWR may be far less valuable than has been claimed.  We encourage the statistical community to develop tools in software packages providing GWR that allow more flexibility in functional form assumptions.

@@

GWR has been presented as a valuable tool for the estimation of site-specific yield response functions used to derive site-specific input rate recommendations from data generated by OFPE [@rakshit2020; @trevisan2021spatial]. @trevisan2021spatial, in particular, reported the economic value of GWR-based variable rate application (VRA) over the optimal uniform rate application (URA) in corn nitrogen management. However, the existing GWR estimations of yield response functions are using a simple linear or quadratic functional form specification. Numerous past agronomic studies recognize that crop yield responses usually follow nonlinear functional forms. For example, corn yield follows a quadratic-plateau response to nitrogen, yet a simple quadratic model was employed by @trevisan2021spatial. Ideally, the GWR model should be specified with quadratic-plateau or more flexible functional forms. However, no software package is readily usable to estimate GWR with quadratic-plateau functional form. The currently used quadratic GWR models are mostly likely mis-specified for yield response functions. Furthermore, GWR itself has faced criticism regarding the statistical reliability of its coefficient estimations [@griffith2008spatial; @harris2019simulation; @paez2011simulation; @da2016multiple]. We hypothesize that the intersection of these two issues—the model mis-specification and the inherent limitations of GWR—may have led to a substantial overestimation of the economic value of the GWR-based VRA over URA. Given the notably high economic value ($58 per hectare) of GWR-based VRA reported in @trevisan2021spatial compared to previous studies on VRA economic evaluation [@bullock1998does; @koch2004economic; @thrikawala1999economic], our suspicion stands a good chance of being plausible.

The primary objective of this study is to examine the degree and direction of bias in the estimation of the economic value of GWR-based variable nitrogen rate recommendations in corn production, especially when the yield response to nitrogen is incorrectly specified as quadratic in the GWR model instead of the correct quadratic-plateau function. (Nathan's emphasis: "a difference between quadratic GWR and a more flexible functional form with no spatial heterogeneity". Shall we put it here?)  We will employ Monte Carlo simulations, which allow us to accurately define and know the true yield response function within our data-generating process, providing a means to quantify the degree of bias present.

Differing from the works of @griffith2008spatial, @harris2019simulation, @paez2011simulation, and @da2016multiple, our goal is not to critique the bias in GWR's coefficient estimates per se. Instead, we seek to assess the economic repercussions^[Note that while the environmental consequences of fertilizer management carry significant societal importance, they fall outside the purview of this study. Herein, "economically optimal" pertains strictly to the maximization of private production profitability for farmers and does not imply "socially optimal."] stemming from functional form mis-specifcatin combined with statistical inaccuracies in the coefficient estimation of the model. We found that a mis-specified GWR model tends to result in substantial overestimation of the economic value of GWR-based VRA over URA.

@@

In practice, since the true responses $f_{j, i}(\cdot)$ are unknown, the calculation of the economic value of the GWR-based VRA is always relying on the assumption that the GWR-estimated yield response functions are true. In this study, the estimated profits of the GWR-based VRA and SCAM-based URA by using the GWR-estimated responses were:

@@

## The true economic value of the GWR-based VRA

The distribution of the true economic value of the quadratic GWR-based VRA over the SCAM-based URA, $\Delta^{GWR}$, from the 1,000 simulations, is shown in Figure \@ref(fig:value-vra). Three different price scenarios are examined to represent the historically low (5.44), middle (6.56), and high (7.67) nitrogen-corn price ratio (\$/kg : \$/kg). Historical corn ($p$) and nitrogen fertilizer ($w$) prices from 2002 to 2022 were obtained from the USDA National Agricultural Statistics Service [@NASS2022] and DTN Retail Fertilizer Trends [@DTN2022], and were transformed into monthly data. The nitrogen-corn price ratios were calculated as $w/p$. Further information regarding the collection, calculation, and summary statistics of the price data can be found in the supplementary appendix (Appendix 1: Historical Corn and Nitrogen Prices).

The true economic values of VRA derived by the mis-specified site-specific response models (quadratic GWRR and GWRT) are only moderate over URA derived by a correctly specified uniform response model (SCAM) when the true yield response function is quadratic-plateau. The median of the true economic values ranges from as small as \$0.81 per hectare to a moderate \$7.83 per hectare, depending on price scenarios and models used. In addition, in any price and model scenario, nearly half of the true economic values are negative, meaning a considerable probability that the GWR-based VRA are actually less profitable than the SCAM-based URA.

The GWRR and GWRT models are slightly different in their economic performances. Specifically, GWRR performs better as nitrogen-corn price ratio increases, while the opposite is the case for GWRT. Figure \@ref(fig:bias-est-eonr) provides insights into this phenomenon, showing the distribution of the average ratio of estimated EONR to true EONR. A value lower than 1 means that EONR is underestimated on average. GWRR tends to overestimate EONR at lower price ratios, but the magnitude of overestimation falls as the price ratio rises. The opposite trend is true for GWRT, where at the high price ratio the degree of EONR overestimation is quite significant. Since a high-price ratio means nitrogen is relatively expensive, the economic cost of overestimation of EONR increases with the price ratio. Figure \@ref(fig:true-vs-estimated-optn-gwr-r) presents for a single simulation round the estimated site-specific EONR plotted against the true site-specific EONR. The site-specific EONR estimates are both inaccurate, but each in its way. The range of EONR estimates by GWRR is very wide because the coefficient on $N^2$ is allowed to vary site specifically in GWRR. Since the estimation of the coefficient on $N^2$ is highly inaccurate, the range of estimated EONRs is much wider than the range of true site-specific EONRs. However, GWRT's range of EONR estimates is very narrow because the coefficient on $N^2$ is constrained to be equal across sites, making GWRT fail to capture the EONR's true spatial heterogeneity.

## Bias in the estimated economic value of the GWR-based VRA

The estimated economic values of the GWR-based VRA over the SCAM-based URA, $\hat{\Delta}^{GWR}$, however, are much larger than the true values. The distribution of the estimated economic values is also shown in Figure \@ref(fig:value-vra), where the median ranges from \$17.15 per hectare to as large as \$45.62 per hectare, surpassing the actual values across all simulation scenarios. Moreover, the estimated economic values are positive in almost all rounds of simulation, suggesting that the GWR-based VRA are almost always more profitable than the SCAM-based URA. But this is significantly erroneous, as nearly half of the time the GWR-based VRA is actually less profitable than SCAM-based URA. These simulation results reveal a substantial overestimation of the GWR-based VRA's economic value when the GWR-estimated site-specific yield response functions are used to substitute the true (but unknown) response functions. That is, in applied research, the quadratic GWR models would tend to create a misleading impression of superior performance in VRA recommendations compared to their actual performance.

The bias in the estimated value, i.e., the discrepancy between the estimated and true economic values $\hat{\Delta}^{GWR} - \Delta^{GWR}$, reaches as large as approximately \$40 per hectare for the GWRR model. When the GWR estimates are considered to define true yield response, the inaccuracy of GWRR generates overestimation of the heterogeneity of yield response within the field, which can lead to significant overestimation of the value of GWRR itself. GWRT exhibits a different pattern of bias. At low price ratios, the GWRT approach displays smaller bias of about \$10 per hectare. This is partially because GWRT tends to understate the heterogeneity of the site-specific yield response functions as was seen in Figure \@ref(fig:true-vs-estimated-optn-gwr-r). As the price ratio rises, bias increases to approximately \$30 per hectare. This is due to the significant overestimation of EONR by GWRT, as illustrated by Figure \@ref(fig:bias-est-eonr). Unlike the GWRR approach, the distribution of the bias by the GWRT approach has a fat tail, indicating that there is a greater chance of encountering large positive bias in the estimated value of VRA over URA.

Figure \@ref(fig:why-bias-many) illustrates how overestimation of the value of VRA can occur. It shows yield response functions estimated by GWRT and the estimated yields associated with the GWR-based EONR rates and SCAM-based EONR rates for 50 randomly chosen cells in a single round of simulation. Since GWR-based EONRs were obtained assuming the yield response functions estimated by GWR, by construction they out-perform the SCAM-based estimates in every cell. For many cells, the SCAM-based estimated EONR is significantly higher than the actual yield-maximizing N rate, resulting in an estimated yield far lower because of the quadratic functional form assumption. Figure \@ref(fig:why-bias-single) panel (a) illustrates the true quadratic-plateau yield response and estimated yield response by the mis-specified quadratic GWR for a single cell. The cell's true yield response plateaus at a nitrogen rate of about $`r round(il_data_oe[aunit_id %in% single_aunit, Nk], digits = 0)`$ kg ha$^{-1}$, but the GWR-estimated yield response function suggests that yield should peak at a nitrogen rate of about $`r round(point_data_gwr[type == "GWR", N], digits = 0)`$ kg ha$^{-1}$ and decline thereafter. The true yield differential between recommended N rates by GWR and SCAM is 0, while the estimated yield differential is about `r round(point_data_gwr[type == "GWR", yield] - point_data_gwr[type == "SCAM", yield], digits = 2)` ton ha$^{-1}$, leading to the overestimation of the GWR yield benefit. Panel (b) illustrates the corresponding true and estimated profit response curves. Based on the true profit response, applying the GWR-based recommendation is about \$`r round(true_yield_data[type != "True", diff(profit)], digits = 0)` ha$^{-1}$ less profitable than the SCAM-based recommendation, where this profit differential comes solely from the difference in nitrogen cost since GWR and SCAM yields are equal. However, when calculations are based on the cell's GWR-estimated response, profit from the GWR-based recommendation exceeds that from the SCAM-based recommendation by about \$`r round(- point_data_gwr[, diff(profit)], digits = 0)` ha$^{-1}$, which is much larger than the true value of \$`r round(true_yield_data[type != "True", diff(profit)], digits = 0)` ha$^{-1}$. Beyond the difference in nitrogen cost, the falsely estimated yield boost by GWR contributes to the profit differential, thus significantly overestimating the actual value of GWR-based VRA.

@@

Our simulation results indicate that variable rate application (VRA) derived from the mis-specified quadratic GWR models (GWRR and GWRT) only has small economic values over uniform rate application (URA) derived from a correctly specified uniform response model (SCAM) when the true yield response function is quadratic-plateau, such as in the corn-nitrogen relationship. However, there is a notable risk of researchers misconstruing the economic values of the quadratic GWR-based VRA as substantially higher than its true values when substituting estimated site-specific yield response functions for the true response functions. This bias in overestimation could be substantial, reaching approximately $40 per hectare. Practitioners in precision agriculture should exercise caution when employing GWR for site-specific input recommendations

Our results clearly show that it can be problematic to use GWR under the assumption of a quadratic functional form when the true yield-input relationship is quadratic-plateau. Previous studies estimating corn yield responses using GWR had to choose the quadratic functional form because no readily usable statistical software or package could be used to implement quadratic-plateau GWR at the time. The only nonlinear model available at the moment is the linear-plateau GWR by @lambert2022geographically. Unfortunately, the estimation of the economically optimal input rate cannot be adjusted according to price using the linear-plateau model as the economically optimal input rate occurs always at the kink (critical input rate where yield plateau is reached). Given that many yield-input relationships follow quadratic-plateau functional form for many crops, it would be fruitful to develop a GWR method with quadratic-plateau functional form. It should be also possible and interesting to build semi-parametric GWR where the yield response functions are estimated without assuming particular functional forms (like SCAM). It is also noted that the recently developed multi-scale GWR model [@fotheringham2017multiscale; @yu2020inference; @comber2022route] allows the bandwidth to be derived separately for each covariate in the model, which would potentially improve the performances of GWR estimates and may need to be incorporated into future research. Note, however, that multi-scale GWR models are not necessary in our simulations as we have only one covariate ($N$) in our simulations.

Most importantly, whether using GWRT or GWRR, researchers must be cautious about their estimated values of VRA when the model is assumed to be quadratic but the true underlying yield-input response is quadratic-plateau. Our results show that it is likely that they will overestimate their true values and the bias can be misleadingly high with non-negligible probability. Indeed, the bias problem we have raised here is likely to be observed for other statistical methods as well because the bias comes from the assumption that the _estimated_ yield response functions are true. Whatever the underlying statistical methods for yield response functions may be, researchers are likely to overestimate the true value of their methods. It would be an interesting future research to investigate the bias in other commonly used statistical methods in the literature on site-specific yield response estimation (such as regression with interaction terms of field characteristics, machine learning, and more) by using a similar simulation-based evaluation approach as done in this study.

@@

```{r, fig.id = "value-vra", fig.cap = "The economic value of the quadratic GWR-based VRA over SCAM-based URA. The true values are calculated using the true yield response functions, while the estimated values are calculated using the quadratic GWR-estimated yield response functions. Two GWR models (GWRR and GWRT) are used. The values are calculated at three nitrogen-corn price ratios ($/kg : $/kg) while the corn price is set at $0.28/kg.", out.width = "7in", fig.width = 6.5}
g_value
```

@@

```{r fig.id = "why-bias-many", fig.cap = "The cause of significant overestimation of the value of GWR-based VRA", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_why_bias_many.png"))
g_why_bias_many
```

```{r fig.id = "why-bias-single", fig.cap = "An illustration of overestimation of the value of GWR-based VRA over SCAM-based URA",fig.width = 6.5, fig.height = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_why_bias_single.png"))
g_why_bias_single
```
